---
title: "Conversion ntsworkflow RDS in JSON"
author: "K Jewell"
date: "2021-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Here we describe the conversion of non-target-screening results from the .report format to JSON format. This description is to help build other data conversion scripts from other non-target-screening software.

The script we are using as an example is called `dbas3.R`. This takes a .report file, e.g. `saale_pos2_i.report`, and a settings file, e.g. `dbas_settings_saale.yaml`, and outputs a json file like `saale_pos2_i.json`. `dbas3.R` is executed using `Rscript` and has not been tested on Windows.

Many of the settings in the .yaml file are only needed for the data evaluation and do not concern the data conversion. So they will not be discussed here.

```{sh}
Rscript dbas3.r dbas_settings_saale.yaml saale_pos2_i.report 
```


# Details

## Structure of .report file

.report files are the output files of a non-target-screening software. It is just an R RC object saved as an RDS file.

```{r}
example <- readRDS("saale_pos2_i.report")
str(example)
```


## Structure of the JSON file

```{json}
{
  "mappings" : {
    "properties" : {
      "ufid" : {"type" : "long"},
      "area" : {"type" : "float"},
      "area_is" : {"type" : "float"},
      "intensity" : {"type" : "float"},
      "intensity_is" : {"type" : "float"},
      "cas" : {"type" : "keyword"},
      "comment" : {"type" : "text"},
      "tag": {"type": "keyword"},
      "data_source" : {"type" : "keyword"},
      "start" : {
        "type" : "date",
        "format" : "yyyy-MM-dd HH:mm:ss||yyyy-MM-dd"
      },
      "km": { "type" : "float" },
      "gkz" : { "type" : "integer" },
      "river" : { "type" : "keyword" },
      "duration" : {"type" : "float"},
      "date_import" : {
        "type" : "date",
        "format" : "epoch_second"
      },
      "eic" : {
        "type" : "nested",
        "properties" : {
          "int" : {
            "type" : "float"
          },
          "time" : {
            "type" : "short"
          }
        }
      },
      "hplc_method" : {"type" : "keyword"},
      "loc" : {"type" : "geo_point"},
      "matrix" : {"type" : "keyword"},
      "ms1" : {                           
        "type" : "nested",                
        "properties" : {
          "int" : {
            "type" : "float"
          },
          "mz" : {
            "type" : "float"
          }
        }
      },
      "ms2" : {
        "type" : "nested",
        "properties" : {
          "int" : {
            "type" : "float"
          },
          "mz" : {
            "type" : "float"
          }
        }
      },
      "mz" : {"type" : "float"},
      "name" : {"type" : "keyword"},
      "norm_a" : {"type" : "float"},
      "pol" : {"type" : "keyword"},
      "rt" : {"type" : "float"},
      "rtt" : {
        "type" : "nested",
        "properties" : {
          "method" : {
            "type" : "keyword"
          },
          "predicted" : {
            "type" : "boolean"
          },
          "rt" : {
            "type" : "float"
          }
        }
      },
      "filename" : {"type" : "keyword"},
      "station" : {"type" : "keyword"}
    }
  }
}
```

See `example_doc.json` or `saale_pos2_i.json`.

## Explanation dbas3.R


```{r, eval=FALSE}
require(shiny)
example$view()  # shiny Fenster zur Visualisierung
```

The necessary data in the .report file are found in the fields `ISresults`, `integRes`, `EIC`, `MS1` and `MS2`

```{r}
head(example$ISresults)
```


```{r}
head(example$integRes)
```

```{r}
head(example$EIC)
```

```{r}
head(example$MS1)
```

```{r}
head(example$MS2)
```

The table `example$integRes` builds the basis, each row will become one first level element in the json array (dbas3.R:51). The other data is added to the `dat` variable which is converted to a `list` on line 199 after which more data is added (done in parallel to speed things up) and then finally outputed to json on line 375.

The settings file (`.yaml`) contains the following fields for the json:
comment, tag, data_source, pol, matrix, duration, station,

In the csv(s) (path in settings file)
gkz, loc, km, river





